---
title: "Investment Committee Memorandum"
subtitle: "Quantitative Equity Portfolio — Photonics & Technology Growth"
author: "Zaeem Hasan | Quant Portfolio Study"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
    latex_engine: xelatex
geometry: margin=1in
fontsize: 11pt
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{xcolor}
  - \usepackage{fancyhdr}
  - \usepackage{graphicx}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{\textit{Confidential — Investment Committee Use Only}}
  - \fancyhead[R]{\thepage}
  - \fancyfoot[C]{\textit{Zaeem Hasan | Quant Portfolio Study}}
  - \definecolor{darkblue}{RGB}{28, 60, 100}
  - \definecolor{lightgray}{RGB}{245, 245, 245}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = 'center',
  fig.pos  = 'H'
)

# ── Dependencies ──────────────────────────────────────────────────────────────
required <- c('tidyverse','knitr','kableExtra','scales','lubridate','ggplot2','gridExtra')
installed <- rownames(installed.packages())
for (pkg in required) {
  if (!pkg %in% installed) install.packages(pkg, repos='https://cran.rstudio.com/')
}

library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(lubridate)
library(ggplot2)
library(gridExtra)

# ── Google Drive path ─────────────────────────────────────────────────────────
# On Posit Cloud: upload CSVs from your Google Drive quant-project folder
# then set this path to wherever they live in your Posit project.
DRIVE_PATH <- './'   # change if files are in a subfolder

# ── Load Module outputs ───────────────────────────────────────────────────────
prices      <- read_csv(paste0(DRIVE_PATH, 'module1_prices.csv'),
                        show_col_types = FALSE) %>%
               rename(date = 1) %>%
               mutate(date = as.Date(date))

returns     <- read_csv(paste0(DRIVE_PATH, 'module1_returns.csv'),
                        show_col_types = FALSE) %>%
               rename(date = 1) %>%
               mutate(date = as.Date(date))

perf        <- read_csv(paste0(DRIVE_PATH, 'module1_per_ticker_perf.csv'),
                        show_col_types = FALSE)

risk        <- read_csv(paste0(DRIVE_PATH, 'module2_risk_summary.csv'),
                        show_col_types = FALSE)

betas       <- read_csv(paste0(DRIVE_PATH, 'module2_betas.csv'),
                        show_col_types = FALSE) %>%
               rename(Ticker = 1)

signals     <- read_csv(paste0(DRIVE_PATH, 'module3_signals.csv'),
                        show_col_types = FALSE)

# ── Helper: pull metric from risk summary ────────────────────────────────────
get_metric <- function(name) {
  as.numeric(risk %>% filter(Metric == name) %>% pull(Value))
}

sharpe      <- get_metric('Sharpe Ratio')
sortino     <- get_metric('Sortino Ratio')
beta_port   <- get_metric('Portfolio Beta')
alpha_ann   <- get_metric('Alpha (ann.)')
ann_ret     <- get_metric('Ann. Return')
ann_vol     <- get_metric('Ann. Volatility')
var_95      <- get_metric('VaR 95% Historical')
cvar_95     <- get_metric('CVaR 95% Historical')

portfolio_value <- perf %>%
  summarise(total = sum(`Current Val ($)`)) %>%
  pull(total)

total_cost <- perf %>%
  summarise(total = sum(`Cost ($)`)) %>%
  pull(total)

total_return_pct <- (portfolio_value / total_cost - 1) * 100
```

\newpage

---

**MEMORANDUM**

**TO:** Investment Committee

**FROM:** Zaeem Hasan, Quantitative Research

**DATE:** `r format(Sys.Date(), '%B %d, %Y')`

**RE:** Quantitative Review — Photonics & Technology Growth Portfolio

**CLASSIFICATION:** Confidential

---

\newpage

# Executive Summary

This memorandum presents a quantitative review of an eleven-position equity portfolio
concentrated in photonics, semiconductor infrastructure, and technology growth.
The portfolio was initiated in August 2025 via two dollar-cost averaged tranches
and has been held without rebalancing through the present date.

**Key findings:**

The portfolio has generated a total return of **`r round(total_return_pct, 1)`%**
since inception against an SPY benchmark return of approximately 5% over the same period,
representing significant alpha generation. Risk-adjusted performance is strong,
with a Sharpe ratio of **`r round(sharpe, 2)`** and a Sortino ratio of
**`r round(sortino, 2)`**, both materially above conventional thresholds.

However, the committee should note three material risks. First, the portfolio
carries a beta of **`r round(beta_port, 2)`** versus SPY, indicating amplified
market sensitivity. Second, two positions — LITE and COHR — account for a
disproportionate share of returns and exhibit a correlation of 0.72, representing
a concentrated single-thesis bet on photonics. Third, the holding period of
approximately eight months is insufficient to draw statistically robust conclusions
about the strategy's long-term viability.

The committee is advised to treat these results as a promising initial signal
warranting continued monitoring rather than a validated long-term strategy.

\newpage

# Portfolio Overview

## Construction

The portfolio was constructed on August 8, 2025 with a second tranche added
August 26, 2025, and a supplemental AMD position added September 13, 2025.
All positions were initiated via fractional share purchases, with approximately
\$3.00 deployed per ticker in the initial tranche.

```{r portfolio-table}
perf %>%
  arrange(desc(`Return (%)`)) %>%
  mutate(
    `Cost ($)`        = dollar(`Cost ($)`, accuracy = 0.01),
    `Entry Price`     = dollar(`Entry Price`, accuracy = 0.01),
    `Current Price`   = dollar(`Current Price`, accuracy = 0.01),
    `Current Val ($)` = dollar(`Current Val ($)`, accuracy = 0.01),
    `P&L ($)`         = dollar(`P&L ($)`, accuracy = 0.01),
    `Return (%)`      = paste0(`Return (%)`, '%')
  ) %>%
  kable(
    booktabs  = TRUE,
    caption   = 'Portfolio Positions — Inception to Present',
    align     = c('l','r','r','r','r','r','r','r')
  ) %>%
  kable_styling(
    latex_options = c('striped','hold_position','scale_down'),
    font_size     = 9
  ) %>%
  row_spec(0, bold = TRUE)
```

## Sector Composition

The portfolio spans five broad themes: photonics (LITE, COHR), semiconductor
infrastructure (AVGO, MRVL, AMD), data center & networking (EQIX, ANET, VRT),
large-cap technology (GOOG), transportation (SNDR), and commodity hedge (GLD).
The photonics and semiconductor positions carry the highest beta and have
been the primary return drivers.

\newpage

# Performance Analysis

## Return Attribution

```{r return-chart, fig.cap="Cumulative portfolio return vs SPY benchmark. Entry August 8, 2025.", fig.height=4}
cum_ret <- returns %>%
  mutate(
    port_cum = cumprod(1 + portfolio) - 1,
    spy_cum  = cumprod(1 + spy) - 1
  )

ggplot(cum_ret, aes(x = date)) +
  geom_line(aes(y = port_cum * 100, colour = 'Portfolio'), linewidth = 1.0) +
  geom_line(aes(y = spy_cum  * 100, colour = 'SPY'),       linewidth = 0.8,
            linetype = 'dashed') +
  geom_hline(yintercept = 0, colour = 'gray60', linewidth = 0.4) +
  scale_colour_manual(values = c('Portfolio' = '#1C3C64', 'SPY' = '#E87040')) +
  scale_y_continuous(labels = function(x) paste0(x, '%')) +
  labs(x = NULL, y = 'Cumulative Return (%)', colour = NULL) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position  = 'bottom',
    panel.grid.minor = element_blank()
  )
```

## Risk-Adjusted Metrics

```{r ratios-table}
tibble(
  Metric = c(
    'Total Return', 'Annualized Return', 'Annualized Volatility',
    'Sharpe Ratio', 'Sortino Ratio', 'Portfolio Beta (vs SPY)',
    'Jensen Alpha (annualized)', 'Max Drawdown (approx.)'
  ),
  Value = c(
    paste0(round(total_return_pct, 1), '%'),
    paste0(round(ann_ret * 100, 1), '%'),
    paste0(round(ann_vol * 100, 1), '%'),
    round(sharpe,    2),
    round(sortino,   2),
    round(beta_port, 2),
    paste0(round(alpha_ann * 100, 1), '%'),
    '-10.4%'
  ),
  Commentary = c(
    'Strong absolute return over 8-month period',
    'Annualized from daily returns; short sample caveat applies',
    'Elevated; consistent with high-beta growth portfolio',
    'Above 2.0 — materially exceeds the conventional 1.0 threshold',
    'Penalizes downside volatility only; higher than Sharpe is favorable',
    'Portfolio amplifies SPY moves by approximately 2x',
    'Return above what beta exposure alone explains',
    'Observed during Nov-Dec 2025 drawdown episode'
  )
) %>%
  kable(
    booktabs = TRUE,
    caption  = 'Risk-Adjusted Performance Summary',
    align    = c('l','r','l')
  ) %>%
  kable_styling(
    latex_options = c('striped','hold_position'),
    font_size     = 9
  ) %>%
  row_spec(0, bold = TRUE)
```

\newpage

# Risk Assessment

## Value at Risk

Three VaR methodologies were applied to the portfolio daily return series.
Historical simulation is the committee's primary reference given the
non-normal return distribution observed (fat left tail relative to parametric fit).

```{r var-table}
tibble(
  Method        = c('Historical Simulation', 'Parametric (Variance-Covariance)', 'Monte Carlo (10,000 simulations)'),
  `VaR 95% (%)` = c(paste0(round(var_95 * 100, 2), '%'), '-3.06%', '-3.08%'),
  `VaR 95% ($)` = c(
    dollar(portfolio_value * var_95,  accuracy = 0.01),
    dollar(portfolio_value * -0.0306, accuracy = 0.01),
    dollar(portfolio_value * -0.0308, accuracy = 0.01)
  ),
  `CVaR 95% (%)` = c(paste0(round(cvar_95 * 100, 2), '%'), '-3.92%', '-3.95%'),
  Note = c(
    'Primary method — no distributional assumption',
    'Underestimates tail risk given observed fat tails',
    'Consistent with parametric; normal distribution assumed'
  )
) %>%
  kable(
    booktabs = TRUE,
    caption  = 'Value at Risk Comparison — Three Methodologies',
    align    = c('l','r','r','r','l')
  ) %>%
  kable_styling(
    latex_options = c('striped','hold_position','scale_down'),
    font_size      = 9
  ) %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(1, bold = TRUE)
```

The historical simulation VaR of `r paste0(round(abs(var_95)*100,2),'%')` exceeds
the parametric estimate, confirming fat tails in the return distribution.
On any given trading day, there is a 5% probability of a portfolio loss exceeding
this threshold. The CVaR of `r paste0(round(abs(cvar_95)*100,2),'%')` represents
the expected loss conditional on breaching that threshold.

## Beta & Concentration Risk

```{r beta-chart, fig.cap="Per-ticker beta versus SPY. Red indicates beta above 1.5.", fig.height=3.5}
betas %>%
  arrange(Beta) %>%
  mutate(
    Ticker = factor(Ticker, levels = Ticker),
    colour = case_when(
      Beta > 1.5 ~ 'High (>1.5)',
      Beta > 1.0 ~ 'Moderate (1.0-1.5)',
      TRUE       ~ 'Low (<1.0)'
    )
  ) %>%
  ggplot(aes(x = Beta, y = Ticker, fill = colour)) +
  geom_col() +
  geom_vline(xintercept = 1.0, linetype = 'dashed', colour = 'gray40', linewidth = 0.6) +
  scale_fill_manual(values = c(
    'High (>1.5)'        = '#e74c3c',
    'Moderate (1.0-1.5)' = '#3498db',
    'Low (<1.0)'         = '#2ecc71'
  )) +
  labs(x = 'Beta vs SPY', y = NULL, fill = 'Beta Category') +
  theme_minimal(base_size = 10) +
  theme(
    legend.position  = 'bottom',
    panel.grid.minor = element_blank()
  )
```

LITE and COHR exhibit the highest betas (~3.0), meaning a 10% decline in SPY
would be expected to produce approximately a 30% decline in these positions
under normal market conditions. GLD (0.53) and EQIX (0.44) provide partial
offset, though their portfolio weight is insufficient to materially dampen
aggregate drawdown.

\newpage

# Signal Analysis

## Current Technical Posture

As of `r format(Sys.Date(), '%B %d, %Y')`, the signal dashboard indicates
a broadly neutral technical environment across the portfolio.

```{r signals-table}
signals %>%
  select(Ticker, Price, RSI, `RSI Signal`, `MA Signal`, `BB Signal`, `Z Signal`, `Composite Score`) %>%
  arrange(desc(`Composite Score`)) %>%
  mutate(Price = dollar(Price, accuracy = 0.01)) %>%
  kable(
    booktabs = TRUE,
    caption  = paste0('Technical Signal Summary — ', format(Sys.Date(), '%B %d, %Y')),
    align    = c('l','r','r','l','l','l','l','r')
  ) %>%
  kable_styling(
    latex_options = c('striped','hold_position','scale_down'),
    font_size     = 9
  ) %>%
  row_spec(0, bold = TRUE)
```

The absence of RSI extremes across all eleven positions is consistent with
a macro volatility compression event (April 2026 tariff environment) rather
than stock-specific exhaustion. AVGO's death cross and MRVL's extended
z-score are the two positions warranting closest near-term monitoring.

\newpage

# Limitations & Study Constraints

The committee should weight the following material limitations when
evaluating the findings of this study:

**Sample size.** The holding period of approximately eight months (165 trading days)
is insufficient to achieve statistical significance on any of the reported ratios.
Sharpe, Sortino, and alpha estimates carry wide confidence intervals at this sample size
and should not be extrapolated as long-term expectations.

**Benchmark selection.** SPY underperformed materially during the study period (~5%),
making outperformance easier to achieve. A fairer benchmark comparison would include
QQQ and a photonics-specific ETF. This will be addressed in v2 of the study.

**Concentration.** LITE and COHR are functionally a single photonics thesis,
correlated at 0.72. The eleven-position count overstates actual diversification.

**Static beta assumption.** Beta was calculated as a single point estimate over
the full holding period. Rolling beta analysis — not yet implemented — would
likely reveal significant regime variation, particularly around the February 2026
drawdown and the April 2026 tariff volatility.

**Portfolio value hardcoded.** The VaR dollar estimates in Module 2 were calculated
against a static portfolio value input rather than the dynamically evolving
portfolio NAV. This will be corrected in v2.

**No transaction cost modeling.** Spreads, platform fees, and tax treatment
are excluded from all return calculations.

\newpage

# Conclusion & Recommendation

The portfolio has generated exceptional risk-adjusted returns over its eight-month
holding period, with a Sharpe ratio of `r round(sharpe, 2)` and annualized alpha
of `r paste0(round(alpha_ann * 100, 1), '%')` above the SPY benchmark.
The photonics thesis — anchored by LITE and COHR — has been the primary return
driver and remains structurally intact based on current technical signals.

**The committee is presented with the following observations:**

1. The risk-adjusted performance merits continued position monitoring
   and potential scale-up subject to position sizing discipline.

2. AVGO warrants a watch flag given its death cross configuration and
   underperformance relative to portfolio peers.

3. MRVL's z-score extension and proximity to the upper Bollinger Band
   suggest limited near-term upside without a fundamental catalyst.

4. The concentration in LITE and COHR represents the single largest
   structural risk in the portfolio. A photonics sector correction
   would be expected to produce outsized drawdown relative to the benchmark.

5. The study's limitations are material and documented. A minimum of
   twelve additional months of data is recommended before drawing
   definitive strategic conclusions.

---

*This memorandum was produced as part of a quantitative research study.
All positions are live holdings. Data sourced from Yahoo Finance via yfinance.
Risk calculations implemented in Python (Modules 1–3) with this memo
compiled in R Markdown (Module 5). Full methodology available in the
accompanying GitHub repository.*

---
